---
title: "HW5"
author: "Adam Ross"
date: "2025-04-15"
output: html_document
---

1. Load house_prices.rda in R...

```{r}
# Load libraries
library(ggplot2)
library(tidyr)
library(dplyr)
library(lubridate)

# Load data
load("I:/.shortcut-targets-by-id/1ehWwunuAo7CE1Vk2JYkUnQMmxh5pph3C/DATA/house_prices.rda")
```

and perform the followings:

(a.)	Use facet_wrap to draw line plots to represent the trend for house_price_index over years for each state. Please have three ticks on the x axis which are for years 1980, 2000, and 2020. Since you have limited space your labels can be ’80, ’00, and ’20 for these ticks.

```{r}
# (1a) Create line plots using facet_wrap
p1 <- ggplot(house_prices, aes(x = date, y = house_price_index)) +
  geom_line() +
  facet_wrap(~state) +
  scale_x_continuous(breaks = ymd(c("1980-01-01", "2000-01-01", "2020-01-01")), labels = c("80", "00", "20"))
  labs(title = "House Price Index Trend by State",
       x = "Year",
       y = "House Price Index")

print(p1)
```

(b.)	Use gather() function from tidyr to reshape your data in a way that it will have measure and value columns for house_price_index and unemploy_perc, respectively, in two columns. You should leave out remaining attributes by using ”-” in front them inside the function. Then, direct the output to a new data frame called house_reshaped Please refer to examples we did in class.

```{r}
# (1b) Use gather() to reshape data
house_prices%>%gather("measure", "value", -state, -house_price_perc, -date)->house_reshaped

head(house_reshaped)
```

(c.)	Use house_reshaped to replicate the graph in 1a with two lines where each represents house_price_index and unemploy_perc over the years for each state. Once you have the chart, please comment on it as to whether it’s a good graph to present those two pieces of information.

```{r}
# (1c) Use house_reshaped to replicate the graph in 1a with two lines
p2 <- ggplot(house_reshaped, aes(x = date, y = value, color=measure)) +
  geom_line() +
  facet_wrap(~state) +
  scale_x_continuous(breaks = ymd(c("1980-01-01", "2000-01-01", "2020-01-01")), labels = c("80", "00", "20"))
  labs(title = "House Price Index Trend by State",
       x = "Year",
       y = "House Price Index")

print(p2)
```

These charts are not a good way to visualize these two data sets because the scale is so different between house_price_idex and unemploy_perc.

(d.) When you achieve the goals above, please create a R markdown file and create an HTML page which has your codes and charts including your comments.